### 2019-10-31 ThesisJo/3_publications.tex
2019-10-22
2019-10-30

%&header
% !TeX root = 2_publications
\input{subheader}
\endofdump

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper I}\label{sec:paperI} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\textbf{An Optimized GPU implementation of the\\
MVDR Beamformer for Active Sonar Imaging}\\
J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen and A. Austeng\\
\textit{IEEE Journal of Oceanic Engineering}\\
Volume 40, Issue 2, April 2015

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper II}\label{sec:paperII} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\textbf{Implementing Capon Beamformer on a GPU for\\ Real-Time Cardiac Ultrasound Imaging}\\
J.P. Åsen, J.I. Buskenes, C.-I.C. Nilsen, A. Austeng and S. Holm\\
\textit{IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control}\\
Volume 61, Issue 1, January 2015

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper III}\label{sec:paperIII} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\textbf{Low Complexity Adaptive Beamforming}~\cite{Buskenes2014}\\
J.I. Buskenes, R.E. Hansen and A. Austeng\\
\textit{IEEE Journal of Oceanic Engineering}\\
Volume 42, Issue 1, January 2017

@@

>>>>

%&header
% !TeX root = thesis
%% !TeX root = 3_publications
\input{subheader}
\endofdump

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper I}\label{sec:paperI} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\textbf{An Optimized GPU implementation of the\\MVDR Beamformer for Active Sonar Imaging}~\cite{Buskenes2014}\\
J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen and A. Austeng\\
\textit{IEEE Journal of Oceanic Engineering}\\
Volume 40, Issue 2, April 2015

\textit{In short}:
\begin{enumerate}
\item Accelerates the MVDR beamformer through algorithmic reductions and fine-tuning to a GPU, yielding a real-time solution.
\item Demonstrates speeds exceeding 1\,Mpixel/s on systems with no more than 32 channels, two orders of magnitude faster than a semi-optimized C-implementation on a CPU.
\item Offers rarely documented insights into hardware limitations, design implications and computational efficiency.
\end{enumerate}

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper II}\label{sec:paperII} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\textbf{Implementing Capon Beamformer on a GPU for\\ Real-Time Cardiac Ultrasound Imaging}~\cite{Asen2013}\\
J.P. Åsen, J.I. Buskenes, C.-I.C. Nilsen, A. Austeng and S. Holm\\
\textit{IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control}\\
Volume 61, Issue 1, January 2015

\textit{In short}:
\begin{enumerate}
\item Alters the implementation in Paper I to operate in beam-space (frequency domain), reduced to its three center beams (the most significant ones).
\item Demonstrates speeds exceeding 1\,Mpixel/s on systems with 64 and 96 channels.
\item Yields similar image quality to element-space MVDR, in accordance with former studies in active sonar.
\end{enumerate}

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper III}\label{sec:paperIII} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\textbf{Low Complexity Adaptive Beamforming}~\cite{Buskenes2016}\\
J.I. Buskenes, R.E. Hansen and A. Austeng\\
\textit{IEEE Journal of Oceanic Engineering}\\
Volume 42, Issue 1, January 2017

\textit{In short}:
\begin{enumerate}
\item Shows that LCA beamformer performs similarly to MVDR with only 6 weight sets to choose from: The rectangular and Kaiser function with $\beta=6$, steered to $\phi\in[0,\pm\frac{1}{2}\delta{\phi_{-3dB}}]$, where $\delta{\phi_{-3dB}}$ is the $-3$\,dB bandwidth of the amplitude response to the weight set.
%Contrasts the LCA and MVDR beamformer, two different ways to solve the minimum variance criterion. The former selects the best performing weights from a predefined set, the latter computes them from a sample covariance matrix estimated from the input data.
\item Finds no further improvement above 15 suitable weight sets.
\item Provides insights into typical amplitude responses of MVDR constrained for active sonar imaging and pinpoints their resemblance to the chosen LCA weight sets. 
\item Suggests swapping the Kaiser function trigonometric one parametrized by $\alpha$, yielding similar image quality while allowing an analytic solution for $\alpha$ to be found. This reduces the algorithm to simply iterating through steering angles.
\end{enumerate}

@@

\textit{In short}:
\begin{enumerate}
\item Implements a SAS simulator on GPU that allows AUVs to compare its images on-the-fly to vast numbers of objects and configurations, yielding superior classification results and real-time responses.
\item Extends a navigational notation system with topological ideas from robotics and homogeneous coordinates, promoting geometrical clarity and validity while aiding the designer to focus on higher level concepts, such as manipulating rigid objects.
\item Demonstrates superior classification accuracy on experimental data of a few thousand objects from Jesus Bay, Norway, at false positive rates (FPR) below 20, similar %, and tied with it above. 
\end{enumerate}

### 2019-10-31 ThesisJo/4_discussion.tex
2019-10-30

\textbf{Paper IV} - \textit{Real-time sonar image synthesis on GPUs for automatic target recognition}:
\begin{enumerate}
\item Implements a SAS simulator on GPU that allows AUVs to compare its images on-the-fly to vast numbers of objects and configurations, yielding superior classification results and real-time responses.
\item Extends a navigational notation system with topological ideas from robotics and homogeneous coordinates, promoting geometrical clarity and validity while aiding the designer to focus on higher level concepts, such as manipulating rigid objects.
\item Demonstrates superior classification accuracy on experimental data of a few thousand objects from Jesus Bay, Norway, at false positive rates (FRP) below 20, similar %, and tied with it above. 
\end{enumerate}

@@

% We grant them perception by performing automatic target recognition (ATR) with an
% acoustic simulator for synthetic aperture sonar (SAS). With it, an AUV can compare
% its images on-the-fly to vast numbers of objects and configurations, yielding superior
% classification results and real-time responses. This is validated on experimental data of
% a few thousand objects from Jesus Bay, Norway, recorded by the Kongsberg Maritime
% HISAS1030 sonar. Our simulator consistently outperformed a conventional method using
% predefined images at false positive rates (FRP) below 20 %, and tied with it above.

>>>>

\textbf{Paper IV} - \textit{Real-time sonar image synthesis on GPUs for automatic target recognition}:
\begin{enumerate}
\item Implements a SAS simulator on GPU that allows AUVs to compare its images on-the-fly to vast numbers of objects and configurations, yielding superior classification results and real-time responses.
\item Extends a navigational notation system with topological ideas from robotics and homogeneous coordinates, promoting geometrical clarity and validity while aiding the designer to focus on higher level concepts, such as manipulating rigid objects.
\item Demonstrates superior classification accuracy on experimental data of a few thousand objects from Jesus Bay, Norway, at false positive rates (FPR) below 20, similar %, and tied with it above. 
\end{enumerate}

@@

% We grant them perception by performing automatic target recognition (ATR) with an
% acoustic simulator for synthetic aperture sonar (SAS). With it, an AUV can compare
% its images on-the-fly to vast numbers of objects and configurations, yielding superior
% classification results and real-time responses. This is validated on experimental data of
% a few thousand objects from Jesus Bay, Norway, recorded by the Kongsberg Maritime
% HISAS1030 sonar. Our simulator consistently outperformed a conventional method using
% predefined images at false positive rates (FPR) below 20 %, and tied with it above.

### 2019-10-31 ThesisJo/background.tex
la2pdf commit - 2015:10:27	11:49:36	(Tue27.Oct)
Getting the hang of including articles
Parallel build starting to work
2019-10-30

@@

Various adaptive beamformers target this limitation by allowing the weights to change for each pixel to better fit the dynamic nature of the incoming wavefield. In other words, they attempt to use the \emph{a priori} information present in the data to improve image quality. The MVDR beamformer is one such method. It finds the set of complex weights that minimizes the beamformer's expected output power, while ensuring unity gain in the look direction~\cite{Capon1969}. This is a convex optimization problem that can be solved using Lagrange multipliers to yield the solution
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a}{\a\T\Ri[n]\a},\label{weights}
\end{gather}
%
where $\a$ is a steering vector and $\R=E\{\x[n]\x\H[n]\}\in\mathbb{C}^{M,M}$ is the spatial covariance matrix for the full array. Since we pre-steer our data to every pixel in the image we simplify (\ref{weights}) by substituting $\a$ with a row vector $\1$ that represents broadside phase-steering. To estimate $\R$ we compute a sample covariance matrix $\eR$. In this computation we perform some degree of:
%
\begin{itemize}
\item \emph{spatial averaging} to avoid signal cancellation by decorrelating coherent echoes~\cite{Kailath1985};
\item \emph{temporal averaging} over an interval comparable to the pulse length (one to five samples) to maintain true speckle statistics~\cite{Synnevag2009a};
\item \emph{diagonal loading} to improve robustness to parameter errors~\cite{Cox1987,Maksym1979}.
\end{itemize}%

@@

>>>>

\input{variables}\ifMonolithic\else\input{header}\input{subheader}\fi
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\providecommand*\Fig{}
\renewcommand*\Fig[1]{Fig.~\ref{#1}}

\providecommand*\a{}
\renewcommand*\a{\vec a}
\providecommand*\k{}
\renewcommand*\k{\vec k}
\providecommand*\p{}
\renewcommand*\p{\vec p}
\providecommand*\R{}
\renewcommand*\R{\mat R}
\providecommand*\Ri{}
\renewcommand*\Ri{\R^{-1}}
\providecommand*\eR{}
\renewcommand*\eR{\mat{\hat R}}
\providecommand*\s{}
\renewcommand*\s{\vec s}
\providecommand*\v{}
\renewcommand*\v{\vec v}
\providecommand*\w{}
\renewcommand*\w{\vec w}
\providecommand*\x{}
\renewcommand*\x{\vec x}

\providecommand*\1{}
\renewcommand*\1{\vec 1}

\providecommand*\H{}
\renewcommand*\H{^{\scriptscriptstyle H}}
\providecommand*\T{}
\renewcommand*\T{^{\scriptscriptstyle T}}

\providecommand*\mat{}
\renewcommand\mat[1]{\boldsymbol{#1}}

\providecommand\bmat{}
\renewcommand\bmat[1]{\begin{bmatrix}#1\end{bmatrix}}
%\newcommand\T{^{\scriptscriptstyle T}}

@@

Various adaptive beamformers target this limitation by allowing the weights to change for each pixel to better fit the dynamic nature of the incoming wavefield. In other words, they attempt to use the \emph{a priori} information present in the data to improve image quality. The MVDR beamformer is one such method. It finds the set of complex weights that minimizes the beamformer's expected output power, while ensuring unity gain in the look direction~\cite{Capon1969}. This is a convex optimization problem that can be solved using Lagrange multipliers to yield the solution
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a}{\a\T\Ri[n]\a},\label{weights}
\end{gather}
%
where $\a$ is a steering vector and $\R=E\{\x[n]\x\H[n]\}\in\mathbb{C}^{M,M}$ is the spatial covariance matrix for the full array. Since we pre-steer our data to every pixel in the image we simplify (\ref{weights}) by substituting $\a$ with a row vector $\1$ that represents broadside phase-steering. To estimate $\R$ we compute a sample covariance matrix $\eR$. In this computation we perform some degree of:
%
\begin{itemize}
\item \emph{spatial averaging} to avoid signal cancellation by decorrelating coherent echoes~\cite{Kailath1985};
\item \emph{temporal averaging} over an interval comparable to the pulse length (one to five samples) to maintain true speckle statistics~\cite{Synnevag2009};
\item \emph{diagonal loading} to improve robustness to parameter errors~\cite{Cox1987,Maksym1979}.
\end{itemize}%

@@

% The ATR algorithms must, however, be sufficiently fast to process the sensor data onboard the AUV in delayed real-time.

% We have developed a simulator for synthetic aperture sonar (SAS) that creates object templates very quickly with the aid of the massive computing power in graphics processing units (GPUs). It loads a 3D model of the sea floor and an object of interest into OpenGL, where emitted sound waves are modeled with a light source placed at the sonar transmit location. We assume rough, isotropic surfaces that reflects sound energy equally in all directions. This can be modeled with a Lambertian scattering model~\cite{Bell1997} where the intensity of the backscattered sound is said to depend only on the on incidence angle of the transmitted signal onto the model surface. When rendering OpenGL is set up to produce an optical 2D intensity image and depth map that reveals the distance from each pixels to the propagation axis. For maximum flexibility this data is finally combined with OpenCL to produce the image templates.
% 
% By coincidence our implementation is not so different from the SIGMAS+ simulator developed at the NATO Undersea Research Centre (NURC)~\cite{Coiras2009a, Coiras2009b}. They both take advantage of OpenGL to render a 2D image from 3D models, are designed for side-looking sonar, and assume a Lambertian scattering model. However, while SIGMAS+ is geared more towards realistically looking sonar images, and include effects such as noise, we do not. Also, their simulator sums images obtained from multiple OpenGL rendering passes to create the sonar image, while we render once and post-process this result with OpenCL to tailor the sonar image full flexibility.

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\ifMonolithic\else\input{subfooter}\fi

### 2019-10-31 ThesisJo/frontmatter.tex
2019-09-30
2019-10-18
2019-10-30

\newpage
\chapter*{Abstract}

\textit{Cognition:}\\The mental action or process of acquiring knowledge and understanding through thought, experience and the senses.

@@

\newcommand\pubEntry[4]{\item #2\\"#1"\\\textit{#3}\\#4}
{\renewcommand{\theenumi}{\textbf{\Roman{enumi}}}
\renewcommand\cleardoublepage{\newpage}
\begin{enumerate}
\pubEntry{%
An Optimized GPU implementation of the MVDR Beamformer for Active Sonar Imaging}%
{J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen and A. Austeng}%
{IEEE Journal of Oceanic Engineering}%
{Volume 40, Issue 2, April 2015}
%
\pubEntry{%
Implementing Capon Beamformer on a GPU for Real-Time Cardiac Ultrasound Imaging}%
{J.P. Åsen, J.I. Buskenes, C.-I.C. Nilsen, A. Austeng and S. Holm}%
{IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control}%
{Volume 61, Issue 1, January 2015}
%
\pubEntry%
{Low Complexity Adaptive Beamforming}%
{J.I. Buskenes, R.E. Hansen and A. Austeng}%
{IEEE Journal of Oceanic Engineering}%
{Volume 42, Issue 1, January 2017}
%
\pubEntry%
{Real-Time Synthetic Aperture Sonar Simulation for Automatic Target Recognition: Notation and Application}%
{J.I. Buskenes, H. Midelfart}%
{IEEE Journal of Oceanic Engineering}%
{Submitted DD.MM.2019}
\end{enumerate}
%}

%\newpage
\chapter*{Excluded Publications}
\addcontentsline{toc}{chapter}{Excluded publications}

@@

>>>>

\newpage
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

\input{abstract}

@@

\newcommand\pubEntry[4]{\item #2\\"#1"\\\textit{#3}\\#4}
{\renewcommand{\theenumi}{\textbf{\Roman{enumi}}}
\renewcommand\cleardoublepage{\newpage}
\begin{enumerate}
\pubEntry{%
An Optimized GPU implementation of the MVDR Beamformer for Active Sonar Imaging}%
{J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen and A. Austeng}%
{IEEE Journal of Oceanic Engineering}%
{Volume 40, Issue 2, April 2015}
%
\pubEntry{%
Implementing Capon Beamformer on a GPU for Real-Time Cardiac Ultrasound Imaging}%
{J.P. Åsen, J.I. Buskenes, C.-I.C. Nilsen, A. Austeng and S. Holm}%
{IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control}%
{Volume 61, Issue 1, January 2015}
%
\pubEntry%
{Low-Complexity Adaptive Sonar Imaging}%
{J.I. Buskenes, R.E. Hansen and A. Austeng}%
{IEEE Journal of Oceanic Engineering}%
{Volume 42, Issue 1, January 2017}
%
\pubEntry%
{Real-Time Synthetic Aperture Sonar Simulation for Automatic Target Recognition: Notation and Application}%
{J.I. Buskenes, H. Midelfart}%
{IEEE Journal of Oceanic Engineering}%
{Submitted DD.MM.2019}
\end{enumerate}
%}

%\newpage
\chapter*{Related Publications}
\addcontentsline{toc}{chapter}{Related publications}

@@

%\newpage
\chapter*{Symbol reference}
\addcontentsline{toc}{chapter}{Symbol reference}

%\begin{table}[H]\noindent%
%\begin{tabular}{>{$}l<{$} >{}l<{} l >{$}l<{$}}
\newcommand\pb[1]{\parbox[c]{9mm}{\textbf{\centering#1}}}
\newcommand\bb[1]{\mathbb{#1}}
%@{\hspace*{0\tabcolsep}}
\begin{longtable}{@{\hspace*{-\tabcolsep}}>{$}l<{$}   >{$\mathbb}l<{$}   >{}l<{} l >{$}l<{$}}
	\pb{Sym-bol}         & \textbf{Space} & \textbf{Unit} & \textbf{Description}                                                    & \textbf{Equal to}                     \\
	\mathcal{A}          &                &               & Name of geometrical object (e.g. an \textbf{A}rray)                     &                                       \\
	\ubar{A}             & \bb V^3\bb A   &               & Frame (configuration) of $\mathcal{A}$                                  & (\uvec{A},\udot{A})                   \\
	\uvec{A}             & \bb V^3        &               & Orientation of $\mathcal{A}$ (absolute, coordinate free)                &                                       \\
	\udot{A}             & \bb A          &               & Position of $\mathcal{A}$ (absolute, coordinate free)                   &                                       \\
	\hat{a}              & \bb V^3        &               & Steering vector (unit-length, coordinate-free)                          &                                       \\
	c                    & \bb R^+        & m/s           & Wave propagation speed                                                  &                                       \\
	\Delta_m             & \bb R          & s             & Delay applied to $m$th sensor                                           &                                       \\
	\dvec{\Delta}        & \bb R^M        & s             & Delay vector, all sensors                                               &                                       \\
	f                    & \bb R          &               & Field strength in (\ref{eq_wave_equation})                              &                                       \\
	                     & \bb R^+        & s$^{-1}$      & Frequency (everywhere else)                                             &                                       \\
	F_s                  & \bb R^+        & s$^{-1}$      & Temporal sampling frequency                                             & T_s^{-1}                              \\
	\vec{k}              & \bb V^3        &               & Wave vector, wavenumber with direction                                  &                                       \\
	\lVert\vec{k}\rVert  & \bb R^+        & m$^{-1}$      & Spatial angular frequency (wavenumber)                                  & \frac{2\pi}{\lambda}=\frac{\omega}{c} \\
	K                    & \bb Z          &               & Number of temporal samples to average (one-sided)                       &  \\
	\lambda              & \bb R^+        & m             & Wavelength                                                              &                                       \\
	L                    & \bb N          &               & Subarray length                                                         &                                       \\
	M                    & \bb N          &               & Number of sensors (hydrophones/microphones)                             &                                       \\
	n                    & \bb Z          &               & Temporal sample index                                                   &                                       \\
	N                    & \bb N          &               & Number of temporal samples                                              &                                       \\
	N_K                  & \bb N          &               & Number of temporal samples to average                                   &                                       \\
	N_L                  & \bb N          &               & Number of subarrays                                                     &                                       \\
	\vec{p}_{AB}         & \bb V^3        &               & Position vector (coordinate-free) from $\udot{A}$ to $\dot{B}$          & \udot{B}-\dot{A}                      \\
	\dvec{p}_{AB}^C      & \bb R^3        &               & Position vector from $\udot{A}$ to $\dot{B}$, represented in $\uvec{C}$ &                                       \\
	t                    & \bb R          & s             & Time                                                                    & T_s n                                 \\
	T_s                  & \bb R          & s             & Temporal sampling period                                                & F_s^{-1}                              \\
	\omega               & \bb R^+        & s$^{-1}$      & Temporal angular frequency                                              & 2\pi f                                \\
	w_m[n]               & \bb C          &               & Weight on $m$th sensor                                                  &                                       \\
	\dvec{w}[n]          & \bb C^M        &               & Weight set, all sensors                                                 &                                       \\
	x_m                  & \bb C^N        &               & Data (delayed, element-space) from $m$th sensor                         &                                       \\
	x_{\textrm{BS},m}    & \bb C^{N_b}    &               & Data (delayed, beam-space) from $m$th sensor                            &                                       \\
	\dvec{x}             & \bb C^{M,N}    &               & Data (delayed, element-space) from all sensors                          &                                       \\
	\dvec{x}_\mathrm{BS} & \bb C^{M,N_b}  &               & Data (delayed, beam-space) from all sensors                             &                                       \\
	z                    & \bb C^N        &               & Beamformer output                                                       &
\end{longtable}
%	y_m                  & \bb C^N        &               & Data (no delay, element-space) from $m$th sensor                        &                                       \\
%\end{tabular}
%\end{table}

%\newpage
\chapter*{Acronym reference}
\addcontentsline{toc}{chapter}{Acronym reference}

\begin{longtable}{l l}
	\textbf{Acronym} & \textbf{Description}                                                      \\
	ADC              & Analog-to-digital converter                                               \\
	ATR              & Automatic target recognition                                              \\
	AUV              & Autonomous underwater vehicle                                             \\
	BS-Capon         & Beam-space Capon (or MVDR, beamformer)                                    \\
	CPU              & Central processing unit                                                   \\
	CUDA             & Compute unified device architecture (from Nvidia)                         \\
	DAC              & Digital-to-analog converter                                               \\
	DAS              & Delay-and-sum (beamformer)                                                \\
	DFT              & Discrete Fourier transform                                                \\
	DMA              & Direct memory access                                                      \\
	ES-Capon         & Element-space Capon (or MVDR, beamformer)                                 \\
	FFT              & Fast Fourier transform                                                    \\
	FPGA             & Field-programmable gate array                                             \\
	FPR              & False positive rate                                                       \\
	FWHM             & Full width at half maximum                                                \\
	GPU              & Graphics processing unit                                                  \\
	HISAS1030        & High resolution synthetic aperture sonar, Kongsberg Maritime (model 1030) \\
	HPC              & High performance computing                                                \\
	HUGIN            & An autonomous underwater vehicle developed by Kongsberg Maritime          \\
	LCA              & Low complexity adaptive (beamformer)                                      \\
   MCM & Mine counter measures \\
	MKL              & Math kernel library (from Intel)                                          \\
	MVDR             & Minimum variance distortionless response (or Capon, beamformer)           \\
	NDC              & Normalized device coordinates                                             \\
	PCI (Express)    & Peripheral component interconnect (data bus)                              \\
	RAM              & Random access memory                                                      \\
	ROC              & Receiver operating curve                                                  \\
	SAS              & Synthetic aperture sonar                                                  \\
	SDFT             & Sliding discrete Fourier transform                                        \\
	SIGMAS+          & Name of an acoustic GPU-based simulator                                   \\
	SSS              & Sidescan sonar                                                            \\
	SM               & Streaming multi-processor \\
   TPR & True positive rate 
\end{longtable}
